---
title: "ASWell_Index"
author: "Julia Gorny"
date: "2023-05-01"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Install once if needed:
# install.packages(c("tidyverse", "writexl", "missRanger", "broom",
#                    "topsis", "factoextra", "DataExplorer"))
library(tidyverse)
library(writexl)
library(missRanger)
library(broom)
library(topsis)
library(ggplot2)
library(factoextra)


set.seed(2023)
```

#Loading the data
Here I am loading the data set. In this dataset I actually deleted the GPI (global peace index) variable

```{r}
df <- read.csv("data_10_variables.csv", header = TRUE, sep = ",") %>%
  select(-any_of(c("X", "X.1")))
```

#Imputing the data
Originally I imputed data using the mean but now I try imputation using missRanger, a random forest imputation which gains more and more popularity recently

```{r}
df_imputed <- missRanger(df, num.trees = 100, verbose = 0)
```

#Getting some insight into the data:
```{r}
DataExplorer::create_report(df_imputed %>% select(OOP:Temp_change)) #this creates different plots for the data (I excluded all character variable to have only the different variables plotted against each other; If I plot this for everything it give some nice insights into the regional distributions though)
```

#Doing the PCA:
Within this step, the data already gets normalised (scaled)
```{r}
pca_data <- df_imputed %>% 
  select(where(is.numeric)) %>% # retain only numeric columns
  select(-Year) %>%  # remove the year variable
  scale() %>% # scale data
  prcomp() # do PCA

# This shows the weightings of the variables 
# for the different PCAs
pca_data$rotation

# This shows the original dataset AND the PCA's 
pca_data %>%
  augment(df_imputed) %>%
  head()

# showing the "importance" of each PC:
pca_data %>%
  tidy(matrix = "eigenvalues")

# This is the scree plot
pca_data %>%
  tidy(matrix = "eigenvalues") %>%
  ggplot(aes(PC, percent)) +
  geom_col() +
  scale_x_continuous(breaks=1:15 ) +
  scale_y_continuous(
    labels = scales::percent_format())  + theme_minimal() 

#this gives an insight into regional distributions
pca_data %>%
  augment(df_imputed) %>% 
  # add original dataset back in
  ggplot(aes(.fittedPC1, .fittedPC2, color=Region)) + 
  geom_point(size = 1.5)

pca_data %>%
  augment(df_imputed) %>% 
  # add original dataset back in
  ggplot(aes(.fittedPC3, .fittedPC4, color=Region)) + 
  geom_point(size = 1.5)

#ordinary biplot
fviz_pca_var(pca_data,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )

```

# normalise the data so that all variables are positive and comparable:

```{r}
min_max_scaled <- df_imputed %>%
  select(where(is.numeric)) %>% # retain only numeric columns
  select(-Year) %>%
  apply(2, function(x) {
    (x - min(x)) / (max(x) - min(x))
  })

#Aggregating the standardised variables into one indicator using the a multi-criteria approach (TOPSIS), within this approach the data gets weighted equally, however directional weighting (impact weighting) will be applied:
impacts <- c(OOP = "-", Water = "+", CPI = "+", partipdem = "+",
             Unemployment = "-", Gini = "-", female_is = "+",
             education = "+", Carbon_T10 = "-", Temp_change = "-")
stopifnot(identical(names(impacts), colnames(min_max_scaled)))


# Calculate the TOPSIS scores with equal weights for all variables and only considering the directional weighting
ASWell_topsis <- topsis(min_max_scaled,
                        weights = rep(1/ncol(min_max_scaled), ncol(min_max_scaled)),
                        impacts = unname(impacts))

#add the my_topsis to the imputed data frame
All_data_ASWell <- cbind(df_imputed, ASWell_topsis)

#download full data
write_csv(All_data_ASWell, 'All_data_ASWell.csv') #this data set shows the data with all the variables and the index scores for each country and yea
```



